An Accurate Model for SARS-CoV-2 pooled RT-PCR Tests Errors

### 2021-03-17 comment.tex
worz

\subsection{RT-PCR}
Before delving into our probabilistic assumptions, we briefly explain
the process of testing for SARS-CoV-2 by RT-PCR. Our exposition
intentionally avoids many details and a comprehensive review of RT-PCR
can be found in \cite{COVID-RTPCR, PCRBook}.

PCR is a process in which a targeted DNA strand's frequency is
amplified to create billions of copies in a reaction mixture. At the
end of this \emph{amplification} process, the presence of targeted DNA
molecules in a reaction mixture can be confidently
determined. Arguably, the best way to understand PCR is as a chain
reaction: Each targeted DNA molecule is copied into two, that are
copied into four and so forth. This cascade gives PCR its name:
polymerase chain reaction.

@@

Since the genetic material of SARS-CoV-2 is RNA rather than DNA, an
extra preprocessing step is required. An enzyme called
reverse transcription (hence the RT in RT-PCR) replicates existing RNA
in the reaction mixture into DNA. Once DNA replicas are made, the PCR
process proceeds as described above, targeting SARS-CoV-2 DNA
replicas.

\section{RT-PCR errors}
Failed detection of SARS-CoV-2 RNA in pooled RT-PCR testing is
referred to below as a false-negative. One possible source of
false-negatives in pooling is sample dilution. When pooling, several
samples are mixed, so the concentration of viral RNA is reduced. This
effect may cause a delay in amplification \cite{DorfmanYuvalDor}, no
detection and, consequently, a false-negative. However, \cite{Lion}
showed that this effect can be safely ignored when mixing up to 32
samples, which we correspondingly assume.

Formally, we consider a pool containing $N$ individuals
$\{1,\dots,N\}$. We denote the true infection states $\theta \in
\{0,1\}^N$, so individual $i$ is infected iff $\theta_i=1$. The RT-PCR
test's sensitivity (true-negative rate) is denoted $\sens$, and the
test's specificity (true-negative rate) is denoted $\spec$. Test results
(data) are denoted $\data$, where $\data=0$ iff the test returned a
positive result.

\subsection{Common assumptions}
Previous studies of group testing strategies assumed that the
false-negative probability does not depend on the number of infected
samples, but merely on the existence of at least one such sample in a
pool \cite{Kim, OptimalDorfmanPool}. Current studies of pooling in the
context of SARS-CoV-2 also employ similar assumptions
\cite{Simplistic1, Simplistic2}. Specifically, these studies assume
that the probability of a negative result is the same for a pool with
a single sample coming from an infected individual and (e.g.)
five. Explicitly:

\begin{align}\label{eq:common}
\Pr(\data=0|\theta) = \left \{
\begin{array}{lr}
        1-\sens & \exists i\ \text{such that } \theta_i = 1\\
        \spec & \text{otherwise}
\end{array} 
\end{align}
We refute the assumptions underlying \eqref{eq:common} in
Section~\ref{subsec:ours} and suggest more plausible ones in
Section~\ref{subsec:ours}.

\subsection{Refuting common assumptions}\label{subsec:refute}
We refute the common assumptions of Section~\ref{subsec:common} with
experimental data summarized in Table \ref{table}, collected from
\cite{Salazar}. There, the authors investigate Dorfman pooling and,
regardless of the pooled test result, follow up and test each pool
member separately. Here, we focus on 128 pools for which at least one
subsequent separate test was positive --- of which 29 pooled tests
were negative and 99 positive. In the data cited in \cite{Salazar}, of
the 29 negative pools, subsequent separate testing yielded a single
positive result in 24. In contrast, of the 99 positive pools, 42
yielded a single positive test upon subsequent separate testing.

The data in Table \ref{table} allows us to test the following null
hypothesis $H_0$: The probability of a pooled false-negative is equal
for pools with one subsequent positively tested member and pools with
two or more such members. We apply Fisher's exact test for the
presence of more than one positive individual in correctly identified
pools. Fisher's test yields an increased odds ratio of 6.4, 95\% CI
(2.2,23.4), with a p-value $\approx 10^{-4}$. Thus we reject $H_0$,
refuting the independence of a pooled result from one or more
individuals infected within the pool, as assumed in \cite{Simplistic1,
  Simplistic2, OptimalDorfmanPool, Kim}

@@

\subsection{Our model}
We assume two pathways for a positive pooled test result: Viral RNA
from an infected individual is correctly detected; or, some erroneous
detection occurs (e.g. contaminant viral RNA is introduced). We assume
a homogeneous and disconnected population (each individual is infected
independently and with equal probability. Erroneous detection of
SARS-CoV-2 RNA (a false-positive) in pooled RT-PCR testing can also
occur. A common assumption \cite{Simplistic1, Simplistic2, Kim,
  OptimalDorfmanPool} is that the false-positive probability does not
depend on the number of negative samples in a pool. We incorporate
this assumption in our likelihood model with a small modification. We
assume that an erroneous amplification can occur in \emph{any}
pool. Specifically, it is possible that correct amplifications fail
and an erroneous one occurs simultaneously. This assumption is
relatively specific for the current application of screening for
SARS-CoV-2 via RT-PCR.  For example, cross-reactivity with other
coronaviruses would have violated this assumption, but it was ruled
out in \cite{KitComparison}.

@@

    \begin{scope}[>={Stealth[black]},
        every node/.style={fill=white},
        every edge/.style={draw=red,very thick}]
      \path [->] (A) edge node {$1-\sens$} (4.5,6.6);
      %\path [->] (B) edge node {$0$} (G);
      %\path [->] (D) edge node {$0$} (G);
      \path [->] (E) edge node {$1-\sens$} (4.5,6.6);
      \path [->] (F) edge node {$\spec$} (5.5,6.6); 
    \end{scope}
  \end{tikzpicture}
  \caption{Illustration of the likelihood model of a pooled test
    result. A pool contains individuals $\{1,2,3,4\}$ with state
    $\theta=(1,0,0,1)$ (i.e. individuals 1 and 4 are infected). A
    negative pooled test implies that three detection paths failed. A
    false negative occurred for $1$ and $4$, each with probability
    $(1-\sens)$. Additionally, no erroneous detection of SARS-CoV-2
    occurred, with probability $\spec$. Individuals $2$ and $3$ are
    not infected and do not contribute to the probability of the
    pooled test result. A positive pooled test arises if any one of
    the above mentioned paths results in a
    detection.}\label{fig:likelihood}
\end{figure}

@@

\section{Application: Calculation of Dorfman's scheme false-negative rate}\label{subsec:detailed}
As an application of our model, we present the differnece in
calculated false-negative rate for a single \emph{infected}
individual, henceforth referred to as "Donald".  We distinguish three
types of false-negative events when performing pooling. A \emph{single
  test's} false-negative is the event of a negative result upon
testing Donald separately, i.e., in an RT-PCR test without pooling. A
\emph{pooled} false-negative occurs when a pooled test containing
Donald's sample (and other samples) yields a negative result, i.e.,
the pooling fails to detect at least one positive result. Lastly, a
\emph{scheme} false-negative occurs when an entire pooling scheme
fails to identify Donald as infected. Our goal is to calculate
Dorfman's scheme false-negative rate. Rephrasing, we wish to answer
the following question: what is the probability of not identifying
Donald as infected under a Dorfman pooling scheme?

Denote the prevalence of infection in the (tested) population $q$. We
give Donald the respect he deserves and assign him as individual $1$,
so that $\theta_1=1$. Then:

\begin{align}
  \begin{split}
    \Pr(\data=0|\theta_1=1) &= \sum_{\theta_2,\dots,\theta_N} \Pr(\theta') \Pr(\data=0|\theta_1=1,\theta_2,\dots,\theta_N) \\
    %
    %
    %
    &= \sum_{k=0}^{N-1}\Pr(\sum_{i=2}^N\theta_i=k)\Pr(\data=0|\theta_1=1,\sum_{i=2}^N\theta_i = k)\\
    %
    %
    %
    &= \sum_{k=0}^{N-1}\binom{N-1}{k}q^k(1-q)^{N-1-k} \Sp(1-\Se)^{1+k}\\
    %
    %
    %
    &= \Sp(1-\Se) \sum_{k=0}^{N-1}\binom{N-1}{k}\left(q(1-\Se)\right)^k(1-q)^{N-1-k}\\
    %
    %
    %
    &= \Sp(1-\Se)(1-q\Se)^{N-1}, \text{ by Newton's Binomial Theorem.} \\
  \end{split}
\end{align}

@@

We calculate the probability that Donald is mistakenly identified as
not infected --- the scheme's false-negative rate --- denoted $\mi$
below. To \emph{correctly} identify Donald as infected, both pooled
and separate tests have to yield a positive result. Thus, the scheme's
false-negative rate $\mi$ is the complement of the product of the two
previous terms:
\begin{align}\label{eq:sfn}
    \begin{split}
        \mi :&= 1 - \Se\Pr(\data=1|\theta=1)\\
        %
        &= 1 - \Se\left [1 - \Sp(1-\Se)(1-q\Se))^{N-1}\right].
    \end{split}
\end{align}

\subsection{Comparison metric}
The single test false-negative rate $1-\Se$ and scheme false-negative
rate $\mi$ are compared via:
\begin{equation}\label{eq:erel}
\rel := \frac{\mi - (1-\Se)}{1-\Se} \cdot 100.
\end{equation}
$\rel$ is the percentage increase in the pooling scheme false-negative
rate, relative to the single test false-negative rate.

The approximation commonly used in the literature estimates the scheme
false-negative rate as $1-\Se^2$ (Section \ref{subsec:assumptions} and
\cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool}). A short
calculation shows that this approximation implies the percentage
increase in scheme false-negative rate is just $100\cdot \Se$.

\subsection{Results}\label{section:results}

We plot $\rel$ for varying prevalence $q$ and sensitivity $\Se$
values. As recommended by \cite{DorfmanYuvalDor}, we apply different
pool sizes $N$, for different prevalence values. We observe that for a
false-positive rate $\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of
reasonable sensitivity and prevalence values \cite{KitComparison,
  InterpretingCOVID19Test, EstimatingRatesLourenco,
  FalsePositiveEstimate}, an increase of at least $60\%$ in $\rel$ can
be expected (Figure \ref{figy}). Interestingly, an increase in
infection prevalence monotonically decreases the scheme false-negative
rate, as can also be easily seen from equation \eqref{eq:sfn}. For the
chosen parameter ranges, the increase in the single test
false-negative rates increases the relative error $\rel$. These
effects can be seen in Figure \ref{figy} (left panel), upon
conditioning on pool size. Extending the range for $\Sp$ yields no
qualitative differences. We further compare $\rel$ to the commonly
used approximation, showing the discrepancy changes as a function of
both prevalence and the single test sensitivity (Figure \ref{figy},
right panel)
\begin{figure}[H]
    \centering
    \includegraphics[width=18cm]{heatmap.jpg}
    \caption{Relative increase in Dorfman pooling false-negative rates
      $\rel$. Left: Colors represent $\rel$, the relative percentage
      increase in the scheme false-negative rates relative to the
      single test false-negative rates (eq \eqref{eq:erel}). Right:
      colors represent the difference between the approximation
      assumed in other studies and $\rel$ as calculated by our
      model. The disease prevalence $q$, is varied on the x-axis,
      while the test sensitivity is varied on the y-axis. Pool size
      $N$, was chosen according to $q$ as in
      \cite{DorfmanYuvalDor}.}\label{figy}
 \end{figure}

\section{Discussion}
Although Dorfman pooling improves testing throughput, we have shown
that it can increase $\rel$ --- the false-negative rates relative to
individual testing. Furthermore, low values of infection prevalence,
or low values of single test false-negative rates, increase
$\rel$. These results remain qualitatively similar under varying
parameter values, in the observed ranges
\cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco,
  InterpretingCOVID19Test} (Figure \ref{figy}).

@@

An approximation to our results has been previously considered: $\mi
\approx 1-\Se^2 $, see \cite{Simplistic1, Simplistic2, Kim,
  OptimalDorfmanPool}. Such an approximation arises when one assumes
that the probability of a false-negative is identical for the pooled
and single tests. Although simplistic in nature, this approximation
does capture the intuition behind our results: For Donald to be
considered negative under Dorfman pooling, he has to test negative
twice. In addition to this approximation only being an upper bound
\cite{Simplistic2}, it does not account for the effect of infection
prevalence on the false-negative rates under Dorfman
pooling. Supporting the importance of this effect, such an association
between prevalence and the scheme false-negative rate under Dorfman
pooling has been empirically noted \cite{DorfmanYuvalDor}.

Although we have shown the inherent risk of Dorfman pooling, this
shortcoming applies to other pooling schemes. Pooling schemes
(e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
positive pooled results to correctly identify Donald as
infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
If the first pool yields a positive result, it is split in two. Then
the splitting is repeated until resulting pools are negative or
individuals are tested separately. With an initial pool size of 32,
Donald will necessarily have to test positive in pools of size 32, 16,
8, 4 and 2, as well as in a single test, for the scheme to correctly
identify him as infected. Compare this to the Dorfman scheme that
requires a positive test in a pool of size $N=8$, and an additional
single positive test to identify Donald as infected. The hierarchical
pooling scheme of \cite{Lion, Kim} will necessarily yield more
false-negatives than Dorfman pooling --- there are additional places
for it to fail.

>>>>

\section{Methods}
Before delving into our probabilistic assumptions, we briefly explain
the process of testing for SARS-CoV-2 by RT-PCR in
Section~\ref{subsec:rtpcr}. Our exposition intentionally avoids many
details and a comprehensive review of RT-PCR can be found in
\cite{COVID-RTPCR, PCRBook}. We then present the \emph{common
  assumption} in Section~\ref{subsec:common} and refute it in
Section~\ref{subsec:refute}. Our model is presented in
Section~\ref{subsec:ours}.

\subsection{RT-PCR}\label{subsec:rtpcr}
PCR is a process in which a targeted DNA strand's frequency is
amplified to create billions of copies in a reaction mixture. At the
end of this \emph{amplification} process, the presence of targeted DNA
molecules in a reaction mixture can be confidently
determined. Arguably, the best way to understand PCR is as a chain
reaction: Each targeted DNA molecule is copied into two, that are
copied into four and so forth. This cascade gives PCR its name:
polymerase chain reaction.

@@

Since the genetic material of SARS-CoV-2 is RNA rather than DNA, an
extra preprocessing step is required. An enzyme called reverse
transcriptase (hence the RT in RT-PCR) replicates existing RNA in the
reaction mixture into DNA. Once DNA replicas are made, the PCR process
proceeds as described above, targeting SARS-CoV-2 DNA replicas.

One possible source of failed detection of SARS-CoV-2 RNA in pooled
RT-PCR testing (a \emph{false-negative}) is sample dilution. When
pooling, several samples are mixed, so the concentration of viral RNA
is reduced. This effect may cause a delay in amplification
\cite{DorfmanYuvalDor}, no detection and, consequently, a
false-negative. Another potential source for false-negatives is a
failed sampling process.

%% However, \cite{Lion} showed that this effect can be
%% safely ignored when mixing up to 32 samples, which we correspondingly
%% assume.

Formally, we consider a pool containing $N$ individuals
$\{1,\dots,N\}$. We denote the true infection states $\theta \in
\{0,1\}^N$, so individual $i$ is infected iff $\theta_i=1$. The RT-PCR
test's sensitivity (true-negative rate) is denoted $\sens$, and the
test's specificity (true-positive rate) is denoted $\spec$. Pooled
test result (data) is denoted $\data$, where $\data=0$ iff the test
returned a negative result.

\subsection{The common assumption}\label{subsec:common}
Previous studies of group testing strategies assumed that the
false-negative probability does not depend on the number of infected
samples, but merely on the existence of at least one such sample in a
pool \cite{Kim, OptimalDorfmanPool}. Current studies of pooling in the
context of SARS-CoV-2 also employ similar assumptions
\cite{Simplistic1, Simplistic2}. Specifically, these studies assume
that the probability of a negative result is the same for a pool with
a single sample coming from an infected individual and (e.g.)
five. Explicitly:

\begin{align}\label{eq:common}
  \Pr(\data=0|\theta) = 
  \begin{cases} 
    1-\sens & \exists i\ \text{such that } \theta_i = 1\\
    \spec & \text{otherwise}
  \end{cases} 
\end{align}

Below, we refer to \eqref{eq:common} as the \emph{common
  assumption}. The essence of the common assumption is that
amplification takes place for all samples at once.
%% We refute the assumptions underlying \eqref{eq:common} in
%% Section~\ref{subsec:ours} and suggest more plausible ones in
%% Section~\ref{subsec:ours}.

\subsection{Refuting common assumption}\label{subsec:refute}
We refute the common assumption with experimental data summarized in
Table \ref{table}, collected from \cite{Salazar}. There, the authors
investigate Dorfman pooling and, regardless of the pooled test result,
follow up and test each pool member separately. We focus on 128 pools
for which at least one subsequent separate test was positive --- of
which 29 pooled tests were negative and 99 positive. In the data cited
in \cite{Salazar}, of the 29 negative pools, subsequent separate
testing yielded a single positive result in 24. In contrast, of the 99
positive pools, 42 yielded a single positive test upon subsequent
separate testing.

The data in Table \ref{table} allows us to test the following null
hypothesis $H_0$: The probability of a pooled false-negative is equal
for pools with one subsequent positively tested member and pools with
two or more such members. $H_0$ is a direct consequence of the
assumptions, and rejecting $H_0$ implies the common assumption are not
realistic, at least for SARS-CoV-2.

We apply Fisher's exact test for the presence of more than one
positive individual in correctly identified pools. Fisher's test
yields an increased odds ratio of 6.4, 95\% CI (2.2,23.4), with a
p-value $\approx 10^{-4}$. Thus, we reject $H_0$, refuting the common
assumption.

@@

\subsection{Our model}
Since the essence of the failed common assumption is that
amplification of all samples occurs only once, we assume the opposite:
amplification of viral RNA occurs for each sample independently.
According to \cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool},
erroneous detection (a false-positive) does not depend on the number
of negative samples in a pool. We incorporate this assumption in our
model with a small modification: We assume that an erroneous
amplification can occur in \emph{any} pool. Specifically, it is
possible that correct amplifications fail and an erroneous one occurs
simultaneously. This assumption is somewhat specific for the current
application of screening for SARS-CoV-2 via RT-PCR. For example,
cross-reactivity with other coronaviruses would have violated this
assumption, but it was ruled out in \cite{KitComparison}.

@@

    \begin{scope}[>={Stealth[black]},
        every node/.style={fill=white},
        every edge/.style={draw=red,very thick}]
      \path [->] (A) edge node {$1-\sens$} (4.5,6.6);
      %\path [->] (B) edge node {$0$} (G);
      %\path [->] (D) edge node {$0$} (G);
      \path [->] (E) edge node {$1-\sens$} (4.5,6.6);
      \path [->] (F) edge node {$\spec$} (5.5,6.6); 
    \end{scope}
  \end{tikzpicture}
  \caption{Illustration of our model. A pool contains individuals
    $\{1,2,3,4\}$ with state $\theta=(1,0,0,1)$ (i.e. individuals 1
    and 4 are infected). A negative pooled test ($\data=0$) implies
    that three detection paths failed. A false-negative occurred for
    $1$ and $4$, each with probability $(1-\sens)$ for
    each. Additionally, no erroneous detection of SARS-CoV-2 occurred,
    with probability $\spec$. Individuals $2$ and $3$ are not infected
    and do not contribute to the probability of the pooled test
    result. A positive pooled test arises if any one of the above
    mentioned paths results in a detection.}\label{fig:likelihood}
\end{figure}

@@

\subsection{Application}
As an application of our model, we calculate the differnece in
false-negative rate for a single \emph{infected} individual,
henceforth referred to as "Donald".  We distinguish three types of
false-negative events when performing pooling. A \emph{single test's}
false-negative is the event of a negative result upon testing Donald
separately, i.e., in an RT-PCR test without pooling. A \emph{pooled}
false-negative occurs when a pooled test containing Donald's sample
(and other samples) yields a negative result, i.e., the pooling fails
to detect at least one positive result. Lastly, a \emph{scheme}
false-negative occurs when an entire pooling scheme fails to identify
Donald as infected. Our goal is to calculate Dorfman's scheme
false-negative rate. Rephrasing, we wish to answer the following
question: what is the probability of not identifying Donald as
infected under a Dorfman pooling scheme?

\subsubsection{Scheme false-negative rate}
Denote the prevalence of infection in the (tested) population $q$. We
give Donald the respect he deserves and assign him as individual $1$,
so that $\theta_1=1$. Then:

\begin{align}
  \begin{split}
    \Pr(\data=0|\theta_1=1) &= \sum_{\theta_2,\dots,\theta_N} \Pr(\theta_2,\dots,\theta_N) \Pr(\data=0|\theta_1=1,\theta_2,\dots,\theta_N) \\
    %
    %
    %
    &= \sum_{k=0}^{N-1}\Pr(\sum_{i=2}^N\theta_i=k)\Pr(\data=0|\theta_1=1,\sum_{i=2}^N\theta_i = k)\\
    %
    %
    %
    &= \sum_{k=0}^{N-1}\binom{N-1}{k}q^k(1-q)^{N-1-k} \Sp(1-\Se)^{1+k}\\
    %
    %
    %
    &= \Sp(1-\Se) \sum_{k=0}^{N-1}\binom{N-1}{k}\left(q(1-\Se)\right)^k(1-q)^{N-1-k}\\
    %
    %
    %
    &= \Sp(1-\Se)(1-q\Se)^{N-1}, \text{ by Newton's Binomial Theorem.} \\
  \end{split}
\end{align}

@@

We calculate the probability that Donald is mistakenly identified as
not infected, henceforth referred to as the scheme's false-negative
rate and denoted $\mi$. In order to correctly identify Donald
as infected, both pooled and separate tests have to yield a positive
result. Thus, the scheme's false-negative rate is:
\begin{align}\label{eq:sfn}
    \begin{split}
        \mi :&= 1 - \Se\Pr(\data=1|\theta=1)\\
        %
        &= 1 - \Se\left [1 - \Sp(1-\Se)(1-q\Se))^{N-1}\right].
    \end{split}
\end{align}

\subsubsection{Comparison metric}
The single test false-negative rate $1-\Se$ and scheme false-negative
rate $\mi$ are compared via:
\begin{equation}\label{eq:erel}
\rel{ours} := \frac{\mi - (1-\Se)}{1-\Se} \cdot 100.
\end{equation}
$\rel{ours}$ is the percentage increase in the pooling scheme false-negative
rate, relative to the single test false-negative rate.

According to the common assumption, the scheme false-negative rate is
$1-\Se^2$. A short calculation shows that this approximation implies
the percentage increase in scheme false-negative rate is just
$\rel{common} := 100\cdot \Se$.

\section{Results}\label{section:results}
We plot $\rel{ours}$ for varying prevalence $q$ and sensitivity $\Se$
values. As recommended by \cite{DorfmanYuvalDor}, we apply different
pool sizes $N$, for different prevalence values. We observe that for a
false-positive rate $\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of
reasonable sensitivity and prevalence values \cite{KitComparison,
  InterpretingCOVID19Test, EstimatingRatesLourenco,
  FalsePositiveEstimate}, an increase of at least $60\%$ in
$\rel{ours}$ can be expected (Figure \ref{figy}). Interestingly, an
increase in infection prevalence monotonically decreases the scheme
false-negative rate, as can also be easily seen from equation
\eqref{eq:sfn}. For the chosen parameter ranges, the increase in the
single test false-negative rates increases the relative error
$\rel{ours}$. These effects can be seen in Figure \ref{figy} (left panel),
upon conditioning on pool size. Extending the range for $\Sp$ yields
no qualitative differences. We further compare $\rel{ours}$ to
$\rel{common}$, showing the discrepancy changes as a function of both
prevalence and the single test sensitivity (Figure \ref{figy}, right
panel).
\begin{figure}[H]
    \centering
    \includegraphics[width=18cm]{heatmap.jpg}
    \caption{Relative increase in Dorfman pooling false-negative rates
      $\rel{ours}$. Left: Colors represent $\rel{ours}$, the relative
      percentage increase in the scheme false-negative rates relative
      to the single test false-negative rates (eq
      \eqref{eq:erel}). Right: colors represent the difference between
      $\rel{common}$ and $\rel{ours}$. The disease prevalence $q$, is
      varied on the x-axis, while the test sensitivity is varied on
      the y-axis. Pool size $N$, was chosen according to $q$ as in
      \cite{DorfmanYuvalDor}.}\label{figy}
 \end{figure}

\section{Discussion}
Although Dorfman pooling improves testing throughput, we have shown
that it can increase $\rel{ours}$ --- the false-negative rates relative to
individual testing. Furthermore, low values of infection prevalence,
or low values of single test false-negative rates, increase
$\rel{ours}$. These results remain qualitatively similar under varying
parameter values, in the observed ranges
\cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco,
  InterpretingCOVID19Test} (Figure \ref{figy}).

@@

%% An approximation to our results has been previously considered: $\mi
%% \approx 1-\Se^2 $, see \cite{Simplistic1, Simplistic2, Kim,
%%   OptimalDorfmanPool}. Such an approximation arises when one assumes
%% that the probability of a false-negative is identical for the pooled
%% and single tests. Although simplistic in nature, this approximation
%% does capture the intuition behind our results: For Donald to be
%% considered negative under Dorfman pooling, he has to test negative
%% twice. In addition to this approximation only being an upper bound
%% \cite{Simplistic2}, it does not account for the effect of infection
%% prevalence on the false-negative rates under Dorfman
%% pooling. Supporting the importance of this effect, such an association
%% between prevalence and the scheme false-negative rate under Dorfman
%% pooling has been empirically noted \cite{DorfmanYuvalDor}.

%% Although we have shown the inherent risk of Dorfman pooling, this
%% shortcoming applies to other pooling schemes. Pooling schemes
%% (e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
%% positive pooled results to correctly identify Donald as
%% infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
%% If the first pool yields a positive result, it is split in two. Then
%% the splitting is repeated until resulting pools are negative or
%% individuals are tested separately. With an initial pool size of 32,
%% Donald will necessarily have to test positive in pools of size 32, 16,
%% 8, 4 and 2, as well as in a single test, for the scheme to correctly
%% identify him as infected. Compare this to the Dorfman scheme that
%% requires a positive test in a pool of size $N=8$, and an additional
%% single positive test to identify Donald as infected. The hierarchical
%% pooling scheme of \cite{Lion, Kim} will necessarily yield more
%% false-negatives than Dorfman pooling --- there are additional places
%% for it to fail.
